---
title: 'R-practice-iris'
---

`iris` data set gives the measurements in centimeters of the variables sepal length, sepal width, petal length and petal width, respectively, for 50 flowers from each of 3 species of iris. The species are Iris setosa, versicolor, and virginica.

Install the `iris` data set on your computer using the command `install.packages("datasets")`. Then load the `datasets` package into your session using the following command.

```{r}
library(datasets)
```

a)  Display the first 6 rows of the `iris` data set

```{r}
# Use head() to get the first 6 rows of the `iris` data set
head(iris)
```

b)  Compute the average of the first four variables (Sepal.Length, Sepal.Width, Petal.Length and Petal.Width) using `sapply()` function.

```{r}
# Use sapply()` function for the average of the first four variable
sapply(iris[1:4], mean, na.rm=T)
```

c)  Show how to use R to replace the missing values in this dataset with plausible ones.

```{r}
# Check the number of missing values in this dataset
sum(is.na(iris)) # 0 missing values
# If the missing values do exist, replace the missing value with 0
iris[is.na(iris)] = 0
# Or we can replace the missing value with the mean, median 
# It depends on if we detect many outliers or not in the dataset
# Checking outliers can be done by plotting
```

d)  Compute the standard deviation for only the first and the fourth variables (Sepal.Length and Petal.Length)

```{r}
# The 4th column should be "Petal.Width"
sapply(iris[,c("Sepal.Length", "Petal.Width")], sd, na.rm=T)
```

e)  Construct a boxplot for `Sepal.Length` and `Sepal.Width` variables, then display the values of all the outliers.

```{r}

# Assign sepalLW to a new data frame with the Sepal.Length, Sepal.Width variables
sepalLW <- data.frame(Sepal.Length = iris$Sepal.Length, Sepal.Width = iris$Sepal.Width)

# Boxplot for `Sepal.Length` and `Sepal.Width`, assigned to myBoxPlot
# Set outline = TRUE to display the values of all the outliers
myBoxPlot <- boxplot(sepalLW,
                     main = "Iris Sepal Length and Sepal Width",
                     outline = TRUE)
# View those outliers
myBoxPlot$out
```

f)  Compute the upper quartile of the `Sepal.Width` variable with two different methods

```{r}
# Method 1: Compute the upper quartile of `Sepal.Width` by summary()
# Use the summary() function on `Sepal.Width` column
# Output is a statistical table
summary(iris$Sepal.Width)
# Extract "3rd Qu." column for the upper quartile
summary(iris$Sepal.Width)["3rd Qu."]

# Method 2: Compute the upper quartile of `Sepal.Width`(Q3, 75% of the data distribution) by the quantile function
quantile(iris$Sepal.Width, .75)
```

g)  Construct a pie chart to describe the species with 'Sepal.Width' less than 4 centimeters.

```{r}
# Filter the iris dataset with 'Sepal.Width' < 4 centimeters
filteredSepalW <- iris[(iris$Sepal.Width < 4), ]
# Use table() function to count the occurences of each factor level of `Species`
countTable <- table(filteredSepalW$Species)
# Make a pie chart of the species with given condition
pie(
  countTable,
  radius = 1,
  main = "Species with Sepal Width less than 4 centimeters")
```
